Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-16 ICML-submission/from-weak-separability-to-strong-separability.tex
Fix beginning of section 5

In this section, we case when the sequence of examples shown by the adversary is
weakly linearly separable
(Definition~\ref{definition:weak-linear-separability}). Throughout this section,
we assume without loss of generality that all examples lie in the unit ball
$\B(0,1) \subseteq \R^d$.\footnote{Consider the following preprocessing step:
upon seeing every example $x_t$, we define $\hat{x}_t = \frac{x_t}{\norm{x_t}}$
as its normalization. It can be easily checked that if $(x_1,y_1), \dots,
(x_T,y_T)$ is weakly linearly separable by margin $\gamma$ and $\norm{x_t} \le
\R$ for all $t$, then the set of normalized examples $(\hat{x}_1,y_1), \dots,
(\hat{x}_T,y_T)$ is weakly linearly separable by margin $\gamma/R$.} Note that
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
alone does not guarantee a finite mistake bound in this setting, as weak linear
separability does not imply strong linear separability.

We consider a kernel function $k(\cdot, \cdot)$ whose corresponding feature map
$\phi(\cdot)$ transforms any sequence of \emph{weakly} linearly separable
examples to \emph{strongly} linearly separable sequence of examples.
Specifically, $\phi$ has the property that if a set of labeled examples in
$B(0,1)$ is weakly linearly separable with a margin $\gamma$, then after
applying $\phi$ the examples become strongly linearly separable with a margin
$\gamma'$ and their norms are bounded by $2$. The parameter $\gamma'$ is a
function of the old margin $\gamma$ and the number of classes $K$, and is
specified in \autoref{theorem:margin-transformation} below.

@@

Consider the classical real separable Hilbert space $\ell_2 = \{ x \in \R^\infty
~:~ \sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped with the standard inner
product $\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$. If we index the
coordinates of $\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of
non-negative integers, the feature map that corresponds to $k$ is $\phi: \B(0,1)
\to \ell_2$,
\begin{align}
& \left(\phi(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \nonumber \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\label{equation:phi}
\end{align}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots,
\alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2!
\dots \alpha_d!}$ is the multinomial coefficient. It can be easily checked that
$$
k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}.
$$
The last equality together with the formula for $k$ implies that $k(x,x) <
+\infty$ for any $x$ in $\B(0,1)$.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ in $\B(0,1) \times
\{1,2,\dots,K\}$ be a sequence of labeled examples that is weakly linearly
separable with margin $\gamma > 0$. Let $\phi$ be as defined in
Equation~\eqref{equation:phi}, and
\begingroup
\allowdisplaybreaks
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(2K-5/2) 2^{K-1}} \\
& \ \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
\end{align*}
\endgroup
Then, the set of labeled examples transformed by $\phi$,
namely $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$, is strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$. In addition, $\norm{\phi(x_1)}_{\ell_2},
\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$.
\end{theorem}

Using this theorem we derive a mistake bound for
Algorithm~\ref{algorithm:kernelized} with kernel function $k(x,x') =
\ip{\phi(x)}{\phi(x')}_{\ell_2}$. This is done by simply invoking
\autoref{theorem:kernelized-upper-bound} with the modified margin $\gamma'$ from
\autoref{theorem:margin-transformation}. The mistake bound is stated as
\autoref{corollary:weakly-separable-examples-mistake-upper-bound} below.

\begin{corollary}[Mistake upper bound]
\label{corollary:weakly-separable-examples-mistake-upper-bound}
Let $K$ be a positive integer and let $\gamma$ be a positive real number. If $L
= ((x_1, y_1), (x_2, y_2), \dots, (x_T, y_T))$ is a sequence of labeled examples
in $V \times \{1,2,\dots,K\}$ such that
\begin{enumerate}
\item $L$ is weakly linearly separable with margin $\gamma$,
\item for all $t$ in $\cbr{1,2,\dots,T}$, $\norm{x_t} \le 1$,
\end{enumerate}
then the expected number of mistakes Algorithm~\ref{algorithm:kernelized} makes
is at most $2^{\widetilde{O}(\min(K \log^2 \frac{1}{\gamma},
\sqrt{\frac{1}{\gamma}} \log K))}$.
\end{corollary}

>>>>

In this section, we consider case when the examples are weakly linearly
separable. Throughout this section, we assume without loss of generality that
all examples lie in the unit ball $\B(0,1) \subseteq \R^d$.\footnote{Instead of
working with feature vector $x_t$ we can work with normalized feature vectors
$\widehat{x}_t = \frac{x_t}{\norm{x_t}}$. It can be easily checked that if
$(x_1,y_1), (x_2,y_2), \dots, (x_T,y_T)$ are weakly linearly separable with
margin $\gamma$ and $\norm{x_t} \le R$ for all $t$, then the normalized examples
$(\widehat{x}_1,y_1), \widehat{x}_2,y_2), \dots, (\widehat{x}_T,y_T)$ are weakly
linearly separable by margin $\gamma/R$.} Note that
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
alone does not guarantee a finite mistake bound in this setting, as weak linear
separability does not imply strong linear separability.

We construct a positive definite kernel function $k(\cdot, \cdot)$ whose
corresponding feature map $\phi(\cdot)$ transforms any sequence of \emph{weakly}
linearly separable examples to \emph{strongly} linearly separable sequence of
examples. Specifically, $\phi$ has the property that if a set of labeled
examples in $\B(0,1)$ is weakly linearly separable with a margin $\gamma$, then
after applying $\phi$ the examples become strongly linearly separable with a
margin $\gamma'$ and their norms are bounded by $2$. The parameter $\gamma'$ is
a function of the old margin $\gamma$ and the number of classes $K$, and is
specified in \autoref{theorem:margin-transformation} below.

@@

Consider the classical real separable Hilbert space $\ell_2 = \{ x \in \R^\infty
~:~ \sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped with the standard inner
product $\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$. If we index the
coordinates of $\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of
non-negative integers, the feature map that corresponds to $k$ is $\phi: \B(0,1)
\to \ell_2$,
\begin{align}
& \left(\phi(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \nonumber \\
& \qquad \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\label{equation:phi}
\end{align}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots,
\alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2!
\dots \alpha_d!}$ is the multinomial coefficient. It can be easily checked that
$$
k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}.
$$
The last equality together with the formula for $k$ implies that $k(x,x) <
+\infty$ for any $x$ in $\B(0,1)$ and thus in particular implies that $\phi(x)$
indeed lies in $\ell_2$.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ in $\B(0,1) \times
\{1,2,\dots,K\}$ be a sequence of labeled examples that is weakly linearly
separable with margin $\gamma > 0$. Let $\phi$ be as defined in
equation~\eqref{equation:phi} and let
\begingroup
\allowdisplaybreaks
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(2K-5/2) 2^{K-1}} \\
& \ \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
\end{align*}
\endgroup
Then, the sequence of labeled examples transformed by $\phi$,
namely $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$, is strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$. In addition, $\norm{\phi(x_1)}_{\ell_2},
\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$.
\end{theorem}

Using this theorem we derive a mistake bound for
Algorithm~\ref{algorithm:kernelized} with kernel function $k(x,x') =
\ip{\phi(x)}{\phi(x')}_{\ell_2}$.

\begin{corollary}[Mistake upper bound]
\label{corollary:weakly-separable-examples-mistake-upper-bound}
Let $K$ be a positive integer and let $\gamma$ be a positive real number. If
$(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \B(0,1) \times \{1,2,\dots,K\}$
is a sequence of weakly separable labeled examples with margin $\gamma > 0$,
then the expected number of mistakes made by Algorithm~\ref{algorithm:kernelized}
with kernel $k(x,x') = \frac{1}{1 - \frac{1}{2}\ip{x}{x'}_{\R^d}}$
is at most $2^{\widetilde{O}(\min(K \log^2 \frac{1}{\gamma},
\sqrt{\frac{1}{\gamma}} \log K))}$.
\end{corollary}

This corollary follows direcly from
Theorems~\ref{theorem:kernelized-upper-bound}
and~\ref{theorem:margin-transformation}.
